---
title: "Connectivity"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  github_document:
    keep_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/02_connectivity/")
```

We are going to use "connectivity" as a measure of conspecific density at each trap location. I'm going to try two different measures of connectivity, __Hanski’s Connectivity Index (CI)__ [(Hanski 1994)](https://www.jstor.org/stable/5591) and the __Conspecific neighbourhood fecundity index (NFI)__ [(Jones & Comita 2008)](https://royalsocietypublishing.org/doi/10.1098/rspb.2008.0894#d3e405). 

They differ in that Hanski's CI includes an exponential function. CI also needs the average dispersal distance of our pre-dispersal seed predators to scale distances, which I'm just going to take as 50m for now.

I had to alter these equations a little. NFI uses the yearly mean total fruit set per tree to weight individuals - we don't have this, so I just used DBH, which assumes that larger individuals produce more fruit [(see here)](https://www.jstor.org/stable/2389015).

__Hanski's Connectivity Index__\

$$CI = \sum exp(-\, \alpha \, dist_{ji}) A_j$$

__Conspecific neighbourhood fecundity index__\

$$NFI = \sum \frac {A_j}{dist_{ij}}$$

_j_ = conspecific tree within a 100 m radius of the trap\
_dist_ = distance between trap i and tree j\
$\alpha$ = 1/ avg. dispersal ability\
_A_ = DBH of tree j\

Note that [LaTeX math won't render in github](https://github.com/rstudio/rmarkdown/issues/806)

```{r, warning=FALSE, message=FALSE}
library("tidyverse"); theme_set(theme_bw(base_size=10))
library("knitr")
library("here")
library("rdist")
library("parallel")

load(here::here("data", "clean", "trapData.RData"))
load(here::here("data", "clean", "treeData.RData"))
trapDat$year <- as.character(trapDat$year)
```

## Calculate euclidean distances
We will create a pairwise matrix with the rdist package to calculate the distance between each tree/trap combo. 

```{r, warning=FALSE}
calculate_dist <- function (species, yr) {
	# subset tree data
	bd <- dplyr::filter(bci, (bci[,"SP4"]==species) & (bci[,"year"]==yr))

	# subset trap data
	td <- dplyr::filter(trapDat, (trapDat[,"SP4"]==species) & 
		(trapDat[,"year"]==yr) )

	# create distance matrix using x and y co-ordinates
	dists <- cdist(bd[,c("gx", "gy")], td[,c("X", "Y")], metric="euclidean")
	dists.df <- as.data.frame(dists)

	# label col names with trap IDs
	colnames(dists.df) <- unlist(td$trap)

	# bind with the bci data
	cbind(bd, dists.df)
}

#for now just apply function to one year and one species
bci_dists <- calculate_dist("CORB", "2015")
```

## Conspecific neighbourhood fecundity index
Lets start by calculating NFI.

```{r, warning=FALSE}
# function to calculate NFI
calculate_NFI <- function (trap) {
  bci_dists %>%
	group_by(treeID, year) %>% 
	mutate(a = dbh / eval(parse(text=trap))) %>% # divide dbh by dist
	ungroup() %>%
	dplyr::select(year, SP4, trap, a) %>%
	group_by(year, SP4) %>%
	summarise(trap = paste(trap), NFI = sum(a), .groups = "drop") # sum over all trees to = 1 value per trap
}

# create list of trap IDs to pass through the function
trap.list <- select(bci_dists,contains("trap_")) %>% colnames()

# apply function to each trap and parallelize
NFIdat <- mclapply(trap.list, calculate_NFI, mc.cores = 4)

# bind output into one dataframe
NFIdat_b <- bind_rows(NFIdat)

head(NFIdat_b) #take a look at NFI values

# merge with trap data
trapConnect <- left_join(NFIdat_b, trapDat,  
	by = c("trap", "year", "SP4"))

```

## Hanski's Connectivity index
Now on to CI using the same format.

```{r, warning=FALSE}
# function to calculate CI
calculate_CI <- function (trap) {
	bci_dists %>%
	group_by(treeID, year) %>% 
	mutate(a = dbh * exp( (-0.02) * eval(parse(text=trap)) ) ) %>% 
	ungroup() %>%
	dplyr::select(year, SP4, trap, a) %>%
	group_by(year, SP4) %>%
	summarise(trap = paste(trap), CI = sum(a), .groups = "drop")
}

# apply function to each trap and parallelize
CIdat <- mclapply(trap.list, calculate_CI, mc.cores = 4)

# bind output into one dataframe
CIdat_b <- bind_rows(CIdat)

head(CIdat_b) #take a look at CI values

```

## Lets see if NFI and CI correlate

```{r CI-NFI, warning=FALSE}
# merge together into one big dataset
trapConnect <- left_join(trapConnect, CIdat_b, by = c("trap", "year", "SP4"))

ggplot(trapConnect) +
	geom_point(aes(x = CI, y = NFI)) +
  ggtitle("Cordia bicolor, 2015")
```

They do, but less so towards larger values which makes sense when you look at the equations. Interesting wedge shape going on. Let's see what the connectivity looks like when plotted against proportion_abscised. I'm going to scale the size of the points with the sum of parts found in the trap.

```{r CORB-2015-connectivity}
ggplot() +
  geom_point(data = subset(trapConnect, SP4 == "CORB"), 
             aes(x = NFI, y = proportion_abscised, size = sum_parts), alpha = 0.5) + 
  ylab("Proportion of seeds abscised") +
  ggtitle("Cordia bicolor, 2015")

ggplot() +
  geom_point(data = subset(trapConnect, SP4 == "CORB"), 
             aes(x = CI, y = proportion_abscised, size = sum_parts), alpha = 0.5) + 
  ylab("Proportion of seeds abscised") +
  ggtitle("Cordia bicolor, 2015")
```

Lots of zeros!
