---
title: "Explore results from models which include a random slope for species"
author: "Eleanor Jackson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    keep_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/06_random-slope-model-results/")
```

In these models (`all_models_quad_rslope`), I have let species have random slopes as well as random intercepts. A random intercept assumes that species are allowed to have their own baseline values (and these are normally distributed around the average intercept). Despite these different starting points, species may not respond to connectivity in exactly the same way - which would need to be accounted for by random slopes... However, this may have made the model too complex, they certainly took a lot longer to run. Let's take a look.

```{r, warning=FALSE, message=FALSE}
library("tidyverse"); theme_set(theme_bw(base_size = 10))
library("broom.mixed")
library("brms")
library("ggdist")
library("loo")
library("patchwork")
options(mc.cores = 4)

# new models
load(here::here("output", "models", "mods_full_quad_rslope.RData")) 
all_models_quad_rslope

# old models
load(here::here("output", "models", "part_pool_quad_models.RData")) 
part_pool_quad_models
```

## Posterior predictive checks (PPC)

```{r dens-pp_checks, fig.height=4, fig.width=7}
pp_check(all_models_quad_rslope$fit[[1]], type = "dens_overlay", nsamples = 50) -> p1
pp_check(all_models_quad_rslope$fit[[2]], type = "dens_overlay", nsamples = 50) -> p2

p1 + p2
```

These look identical to the pp_check plots from the models without random slopes.

## Compare the predictive accuracy of the models using LOO-CV

```{r loo}
comp <- loo_compare(part_pool_quad_models$fit[[1]], part_pool_quad_models$fit[[2]],
                    all_models_quad_rslope$fit[[1]], all_models_quad_rslope$fit[[2]])

print(comp, digits = 3)
```

Interesting! loo says from best to worst:

* random slope ZOIB
* non-random slope ZOIB
* random slope ZIB
* non-random slope ZIB

## Model Results
Compare results from these new models to the non-random slope models.

```{r compare-results}
part_pool_quad_models %>%
  mutate(posterior = map(fit, ~brms::posterior_samples(.x, pars = c("b_CI_pred.sc")))) %>%
  select(model, posterior) %>%
  unnest(posterior) %>% 
  pivot_longer(b_CI_pred.sc, names_to="mu" ) %>%
  ggplot(aes(x = as.factor(model), y = as.numeric(value))) +
  coord_flip() +
  labs(y = "Estimate ± CI [95%]", x = "") +
  scale_y_continuous(limits = c(-1, 1)) +
  ggdist::stat_halfeye(.width = c(.90, .5), normalize = "xy", limits = c(-3, 3)) +
  geom_hline(yintercept = 0, linetype = 2, size = 0.25) +
  ggtitle("(1 | trap) + (1 | year) + (1 | quadrat) +\n(1 | species) ")-> p3

all_models_quad_rslope %>%
  mutate(posterior = map(fit, ~brms::posterior_samples(.x, pars = c("b_CI_pred.sc")))) %>%
  select(model, posterior) %>%
  unnest(posterior) %>%
  pivot_longer(b_CI_pred.sc, names_to="mu" ) %>% 
  ggplot(aes(x = as.factor(model), y = as.numeric(value))) +
  coord_flip() +
  labs(y = "Estimate ± CI [95%]", x = "") +
  scale_y_continuous(limits = c(-1, 1)) +
  ggdist::stat_halfeye(.width = c(.90, .5), normalize = "xy", limits = c(-3, 3)) +
  geom_hline(yintercept = 0, linetype = 2, size = 0.25)  +
  ggtitle("(1 | trap) + (1 | year) + (1 | quadrat) +\n(1 + connectivity | species)") -> p4

p3 + p4
```

The posterior distribution is much more spread out now! I think this was to be expected as we are allowing more variation. I think it's good that the estimates look very similar in the different approaches, suggests that they are quite robust?

Let's take a quick look at the species random effects from the new ZOIB.

```{r species-re, fig.height=8, fig.width=5}
all_models_quad_rslope %>%
  filter(model == "ZOIB") %>%
  mutate(posterior = map(fit, ~brms::posterior_samples(.x))) %>%
  select(model, posterior) %>%
  unnest(posterior) %>% 
  select(contains("r_SP4[")) %>% 
  select(contains(",Intercept]")) %>% 
  pivot_longer(cols = contains("r_SP4["), names_to = "SP4") %>%
  mutate(SP4 = as.character(SP4)) %>%
  mutate(SP4 = gsub("r_SP4[", "", SP4, fixed = TRUE)) %>%
  mutate(SP4 = gsub(",Intercept]", "", SP4, fixed = TRUE))-> rslope_ZOIB_plot_dat

ggplot(rslope_ZOIB_plot_dat, aes(x = reorder(SP4, value), y = as.numeric(value))) +
  ggdist::stat_halfeye(.width = c(.90, .5), normalize = "xy", limits = c(-3, 3), size = 1.5) +
  coord_flip() +
  labs(y = "Estimate ± CI [95%]", x = "") +
  geom_hline(yintercept = 0, linetype = 2, size = 0.25)
```

From a quick eyeball they don't look different from the old models, the species seem to be in the same order.

## To conclude..
These models look good! I think that the random slope is good in both a model-fitting sense and an ecological sense.

However, note that I did get some warnings when the ZIB was fit: 
"1: Bulk Effective Samples Size (ESS) is too low, indicating posterior means and medians may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#bulk-ess 
2: Tail Effective Samples Size (ESS) is too low, indicating posterior variances and tail quantiles may be unreliable.
Running the chains for more iterations may help. See
http://mc-stan.org/misc/warnings.html#tail-ess "

I did not get these warnings with the ZOIB. Stan recommends running for more iterations - I will try this if I decide to take the ZIB model further, but perhaps now is the time to drop it. The ZOIB has performed better in every iteration so far.

## Try nesting trap inside quadrat, added 21-06-07
I have run a model which has nested random effects: intercept varying among quadrats and among traps within quadrats.

```{r nest-result, fig.height=4, fig.width=10}
quad_rslope_nest <- readRDS(here::here("output", "models", "full_sp_quad_rslope_nest_ZOIB.rds"))

brms::posterior_samples(quad_rslope_nest, pars = c("b_CI_pred.sc")) %>% 
  pivot_longer(b_CI_pred.sc, names_to="model") %>%
  mutate(model = recode(model, b_CI_pred.sc = "ZOIB")) %>%
  ggplot(aes(x = as.factor(model), y = as.numeric(value))) +
  coord_flip() +
  labs(y = "Estimate ± CI [95%]", x = "") +
  scale_y_continuous(limits = c(-1, 1)) +
  ggdist::stat_halfeye(.width = c(.90, .5), normalize = "xy", limits = c(-3, 3)) +
  geom_hline(yintercept = 0, linetype = 2, size = 0.25)  +
  ggtitle("(1 | quadrat/trap) + (1 | year) +\n(1 + connectivity | species)") -> p5

p3 + p4 + p5
```

Estimate looks similar to the non-nested ZOIB!

```{r loo2}
comp2 <- loo_compare(part_pool_quad_models$fit[[2]], all_models_quad_rslope$fit[[2]], quad_rslope_nest)

print(comp2, digits = 3)
```

Nesting trap within quadrat has not improved the model fit compared to the random slope ZOIB. `quad_rslope_nest` is very slightly worse.
