---
title: "Estimate fruit set using flowers"
author: "Eleanor Jackson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    keep_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/09_estimate-fruit-set/")
```

Can/should we include analogous analyses on fruit set (proportion of flowers turning into fruits) for some of the species?

```{r, warning=FALSE, message=FALSE}
library("tidyverse"); theme_set(theme_bw(base_size = 10))
load(here::here("data", "clean", "trapData.RData"))
```

Part 6 is perfect and female flowers, part 9 is male flowers. We are only interested in female flowers and can use these as an estimate of abscised flowers, successful female flowers will have turned into fruits, so we can do:

(abscised flowers * nseedfull) / total seeds

N.B.
The variable ‘quantity’ was recorded as presence only (i.e., all values equal 1) for flowers (part=6 or part=9) and buds (part=0) until 20020709. Starting on 20020709, the variable quantity was recorded on an approximate logarithmic scale (i.e., 1, 10, 100 or 1000) for flowers and buds.

## Calculate the proportion of ovules that were prematurely abscised

```{r}
# seed rain data
seedRain <- read.table(here::here("data", "raw",
	"BCI_TRAP200_20190215_spcorrected.txt"),
	header=TRUE, stringsAsFactors = FALSE)

seedTrait <- read.csv(here::here("data", "raw",
	"20120227_seedsMassForTraits.csv"),
	header=TRUE, stringsAsFactors = FALSE)

seedRain <- rename(seedRain, SP4 = sp)

seedTrait <- seedTrait %>%
   select(LIFEFORM, N_SEEDFULL, SP4, GENUS, SPECIES)

seedRain$fecha <- as.character(seedRain$fecha)
seedRain$fecha <- as.Date(seedRain$fecha, "%Y-%m-%d")
seedRain$year <- format(as.Date(seedRain$fecha), "%Y")

seedDat <- left_join(seedRain, seedTrait, by = c("SP4"))

seedDat <- subset(seedDat, LIFEFORM != "EPIPHYTE"|LIFEFORM != "HEMIEPIPHYTE"|LIFEFORM != "HERB"|LIFEFORM != "VINE")

seedDat <- seedDat %>%
	drop_na("N_SEEDFULL")

# don't have the full data for these years
seedDat %>%
  filter(year >= 2003 & year != 2019) -> seedDat

# format trap id
seedDat$trap <- formatC(seedDat$trap, width = 3, format = "d", flag = "0")
seedDat$trap <- paste("trap", seedDat$trap, sep="_")

seedDat %>%
	group_by(part, SP4, year, trap, N_SEEDFULL, GENUS, SPECIES) %>%
	summarise(quantity_sum = sum(quantity, na.rm = TRUE)) %>%
	ungroup() -> absdat

absdat %>% 
  subset(part == 6) %>%
	replace(., is.na(.), 0) %>%
	rowwise() %>%
	mutate(abscised_ovules = quantity_sum*N_SEEDFULL) %>%
	ungroup() %>% 
	group_by(SP4, year, trap, GENUS, SPECIES) %>%
	summarise(abscised_ovules = sum(abscised_ovules, na.rm = TRUE)) %>%
	ungroup() -> flower_dat 

glimpse(flower_dat)

# join up with fruit abscission data and calculate the proportion of ovules abscised
left_join(trapDat, flower_dat, by = c("SP4", "year", "trap", "GENUS", "SPECIES")) %>%
  mutate(proportion_o_abscised = abscised_ovules / total_seeds) -> trapDat_fs
  
```

## can now take a look at it

```{r}

trapDat_fs %>%
  filter(SP4 == "HYEL" & year >= 2003) %>%
  ggplot(aes(x = year, y = proportion_o_abscised, colour = trap)) +
  geom_point() +
  theme(legend.position = "none")

```

Can't really say much from this brief foray into fruit set, but I think we aren't going to go down this route for the current manuscript. We have run out of time and there are some unknowns about how reliably flowers will fall into the seed traps.
