---
title: "Using capsules to estimate fruit drop for vertebrate dispersed species"
author: "Eleanor Jackson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    keep_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/10_using-capsules/")
ggplot2::theme_set(ggplot2::theme_bw(base_size = 10))
```

For the analysis thus far we have assumed that the probability of a fruit or seed falling into a trap is the same for mature and immature fruits and seeds. However, we have concerns about including species for which the ratio of immature to mature fruits could be affected by removal of mature fruits from the canopy. (If animals forage for fruit in a density-dependent fashion, this could create patterns in the seed trap data that are hard to interpret.)
 
To avoid those complications, we were thinking that it might be worth focusing on species which are not animal dispersed (and therefore presumably unlikely to be routinely eaten by animals). SJW suggested that for some species it is possible to reconstruct the numbers of mature fruits based on capsules and similar evidence found in the traps.

Capsules are recorded as `part 3` in the seed rain dataset. *"Capsules - This is a part that vertebrates never eat. Botanically this might be a capsule, pedicel, bract, etc.* SJW and OC also record *"Fragments of fruit dropped by vertebrates (we record the number of fruit represented by counting pedicels or the points of attachment of fruits to the mother plant)."* these are `part 4` in the dataset.

In [Gripenberg et al. 2019](https://doi.org/10.1111/ele.13359) they use capsules to estimate seed abundances. In the supplementary material they write:
*"To estimate the total number of seeds produced for each species, we multiplied the number of fruits collected by the species-specific average number of seeds per fruit, and then added either the number of seeds or the number of capsules multiplied by species-specific average number of fruits per capsule multiplied by average number of seeds per fruit; whichever number was larger. The rationale for this approach is that since some of the seeds falling into the trap might have originated from the capsule, we might overestimate seed numbers if we were to use (fruits × average seeds/fruit) + seeds + (capsules × average seeds/fruit). Since many seed predators are likely to be pre-dispersal seed predators attacking the fruits before they reach maturity, immature fruits were included in the estimates."*

Here, I am going to use capsules to estimate fruit drop and compare this against the values we estimated using `part 1` - mature fruits.

notes:

* could multiple capsule parts have come off of a single fruit?
* could capsule parts have come from immature fruits?
* do we have average number of fruits per capsule? - Yes

SJW: *"If you need a few examples to get started on this, use any or all species of Trichilia, Virola, Protium, Tetragastris (which is now in Protium isthmensis [spelling?])."*

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("patchwork")
library("ggforce")

# seed rain data
read.table(
  here::here("data", "raw", "BCI_TRAP200_20190215_spcorrected.txt"),
  header = TRUE,
  stringsAsFactors = FALSE
) -> seed_rain

# seed traits from joe
read.csv(
  here::here("data", "raw", "20120227_seedsMassForTraits.csv"),
  header = TRUE,
  stringsAsFactors = FALSE
) %>%
  select(LIFEFORM, N_SEEDFULL, SP4) -> seed_trait

```

### Clean up seed rain data

```{r}
seed_rain %>%
  mutate(fecha = as.character(fecha)) %>%
  mutate(fecha = as.Date(fecha, "%Y-%m-%d"),
         year = format(as.Date(fecha), "%Y")) %>%
  left_join(seed_trait, by = c("sp" = "SP4")) %>%
  
  # don't have the full data for these years
  filter(year != "1987" & year != "2019") %>%
  
  # keep only woody plants
  filter(
    LIFEFORM != "EPIPHYTE" |
      LIFEFORM != "HEMIEPIPHYTE" |
      LIFEFORM != "HERB" | LIFEFORM != "VINE"
  ) %>%
  drop_na("N_SEEDFULL") %>%
  mutate(sp = tolower(sp),
         trap = formatC(
           trap,
           width = 3,
           format = "d",
           flag = "0"
         )) %>%
  mutate(trap = paste("trap", trap, sep = "_")) %>%
  rename(sp4 = sp, n_seedfull = N_SEEDFULL) -> seed_dat

# sum weekly counts of parts into yearly counts per trap per species
seed_dat %>%
  filter(part == 1 | part == 2 | part == 5 | part == 3 | part == 7) %>%
  group_by(part, sp4, year, trap, n_seedfull) %>%
  summarise(quantity_sum = sum(quantity, na.rm = TRUE),
            .groups = "drop") -> abs_dat

# combine with species data to drop species that we can't use
read.csv(here::here("data", "clean", "species_list.csv")) %>%
  filter(dioecious == FALSE & ((animal_disp == FALSE | is.na(animal_disp)) | capsules == TRUE)) %>%
  select("sp4", "capsules", "capsules_per_fruit") %>%
  inner_join(abs_dat, by = "sp4") -> abs_dat_comb

```

### Calculate viable seeds

if capsule = TRUE,  viable seeds = (mature fruits * avg seeds per fruit) + 
  ( (capsules * avg seeds per fruit) __OR__ single diaspores)

part 1 + part 2 __OR__ part 1 + part 3

if capsule = FALSE, viable seeds = (mature fruits * avg seeds per fruit) + 
  single diaspores

part 1 + part 2

```{r}
abs_dat_comb %>%
  subset(part == 1 | part == 2 | part == 3) %>%
  
	rowwise() %>%
	mutate(parts_avgseeds = case_when(
	  
	   # mature fruit
	  part == 1 ~ quantity_sum * n_seedfull,
	  
	  # single diaspores
	  part == 2 ~ quantity_sum, 
	  
	  # capsules
	  part == 3 ~ (quantity_sum / capsules_per_fruit) * n_seedfull)) %>%
  
  # estimate number of mature fruits based on fruits + capsules, and based on fruits + seeds
  group_by(sp4, year, trap, capsules) %>%
  summarise(capsules_seeds = ifelse(
	  part == 1 | part == 3, sum(parts_avgseeds, na.rm = TRUE), 0), 
	  
         fruits_seeds = ifelse(
	  part == 1 | part == 2, sum(parts_avgseeds, na.rm = TRUE), 0),
	  .groups = "drop") %>%
  
  # if fruits + seeds estimate is larger, use that rather than fruits + capsules
  mutate(fruits_capsules = 
           ifelse(fruits_seeds > capsules_seeds, fruits_seeds, capsules_seeds)) %>%
  
  # if we can't estimate fruits from capsules for a species, use fruits + seeds
  mutate(viable_seeds = case_when(
	  capsules == FALSE | is.na(capsules) ~ fruits_seeds,
	  capsules == TRUE ~ capsules_seeds)) %>%
  
  select(-capsules_seeds, -fruits_seeds) -> abs_dat_viable
  
```

### Calculate abscised seeds

```{r}
# abscised seeds,  part 5 = immature fruit 
abs_dat_comb %>%  
  subset(part == 5) %>%
	rowwise() %>%
	mutate(abscised_seeds = quantity_sum*n_seedfull) %>%
	group_by(sp4, year, trap) %>%
	summarise(abscised_seeds = sum(abscised_seeds, na.rm = TRUE), 
	          .groups = "drop") -> abs_dat_abscised

```

### Calculate proportion abscised

```{r}
# join abscised and viable seeds data, where there is no match will = NA, change to zero
full_join(abs_dat_abscised, abs_dat_viable, by= c("sp4", "year", "trap")) %>%
  mutate_at(vars(viable_seeds, abscised_seeds), ~replace(., is.na(.), 0)) -> abs_dat_abscised_viable

# calculate proportion abscised
abs_dat_abscised_viable %>%
	rowwise() %>%
	mutate(total_seeds = sum(abscised_seeds, viable_seeds, na.rm = TRUE), 
	       proportion_abscised = abscised_seeds / total_seeds) -> prop_dat

abs_dat %>%
	group_by(sp4, year, trap) %>%
	summarise(sum_parts = sum(quantity_sum, na.rm = TRUE), .groups = "drop") -> sum_dat

prop_dat %>%
	left_join(sum_dat, by = c("sp4", "year", "trap")) -> trap_dat_caps

```

# Compare with old (non-capsule) data

First load in the original `proportion_abscised` values and join with the new dataset that uses capsules to estimate viable seeds.
```{r}
load(here::here("data", "clean", "trapData.RData"))

trapDat %>%
  select(SP4, year, trap, proportion_abscised) %>%
  mutate(SP4 = tolower(SP4)) -> trap_dat_select

trap_dat_caps %>% 
  select(sp4, year, trap, proportion_abscised, capsules, total_seeds) %>%
  rename(proportion_abscised_caps = proportion_abscised) %>%
  full_join(trap_dat_select, by = c("sp4" = "SP4", "year", "trap")) -> comp_dat

```

```{r warning=FALSE, compare-capsule-method}
comp_dat %>%
  filter(capsules == TRUE) %>%
  distinct(sp4) -> sp4_caps_true
  
comp_dat %>% 
  filter(sp4 %in% sp4_caps_true$sp4 & total_seeds >= 5) %>%
  pivot_longer(cols = c(proportion_abscised_caps, proportion_abscised), names_to = "method") %>%
  ggplot(aes(x = value, colour = method)) +
  geom_histogram(fill = "white", position="dodge") +
  ggforce::facet_wrap_paginate(~sp4, scales = "free_y",  ncol = 4, nrow = 4) +
  theme(legend.position = "top") -> p
    
ggforce::n_pages(p) -> n_pages

for(i in 1:n_pages){
    print(p + ggforce::facet_wrap_paginate(~sp4, scales = "free_y",  ncol = 4, nrow = 4, page = i))
}

```

For many species there is little difference between the proportion abscised calculated using the capsules method and the original method. If the original estimate of viable seeds was larger than the estimate calculated using capsules, the original estimate was used. This suggests that few capsules were collected for most species.

# Fruits with insect emergence holes

SJW: *"I believe I have a Christmas present for you. Check out the definition of part=7 in the metadata for the 200 trap study. There are a handful of species whose fruits are abscised with conspicuous “weirdnesses” when infested by an insect. Faramea occidentalis is an example that Sofia will know well. For this handful of species, we know an insect predator is the cause of immature fruit fall. For other species, you only have the reasonable inference that enemies cause immature fruit fall. Results for the handful of species with part=7 will, hopefully, substantiate this inference."*

I‘ll have a look at the data for species with part 7 counts to see how many fruits were dropped immaturely without insect emergence holes VS. fruits with insect emergence holes. I think Joe's idea was that if the proportion of part 7 looked similar to the proportion of part 5 (immature fruits) then it will back-up our assumption that seed enemies are causing premature fruit drop. 

We aren't actually going to be using _Faramea occidentalis_ in the model because it's animal dispersed and it's seeds cannot be estimated from capsules.

How many other species have a part 7?

```{r paged.print=TRUE}

abs_dat_comb %>%
  filter(part == 7) %>%
  group_by(sp4) %>%
  summarise(
    sum_part_7 = sum(quantity_sum)
  ) %>%
  arrange(desc(sum_part_7))

```

79 is more species than I was expecting. Let's look at `thim` (*Thinouia myriantha*) as it has the highest count of part 7.

```{r, fruits-with-insect-holes}
# calculate proportion part 7
abs_dat_comb %>%
  filter(sp4 == "thim" & (part == 7 | part == 1 | part == 2)) %>%
  rowwise() %>%
  mutate(
    parts_avgseeds = case_when(
      # mature fruit
      part == 1 ~ quantity_sum * n_seedfull,
      
      # single diaspores
      part == 2 ~ quantity_sum,
      
      # fruit with insect emergence holes
      part == 7 ~ quantity_sum * n_seedfull
    )
  ) %>%
  group_by(sp4, year, trap) %>%
  summarise(
    insect_abscised_seeds = sum(parts_avgseeds[part == 7]),
    viable_seeds = sum(parts_avgseeds[part == 1 | part == 2]),
    .groups = "drop"
  ) %>%
  group_by(sp4, year, trap) %>%
  mutate(
    total_seeds = sum(insect_abscised_seeds, viable_seeds, na.rm = TRUE),
    proportion_insect_abscised = insect_abscised_seeds / total_seeds
  )  -> thim_part7


trap_dat_caps %>%
  filter(sp4 == "thim") %>% 
  filter(total_seeds >=5) %>%
  ggplot(aes(x = proportion_abscised)) +
  geom_histogram() +
  xlim(0, 1) +
  ylim(0, 200) +
  ggtitle("Proportion part = 5\nimmature fruits") -> p1

thim_part7 %>% 
  filter(total_seeds >=5) %>%
  ggplot(aes(x = proportion_insect_abscised)) +
  geom_histogram() +
  xlim(0, 1) +
  ylim(0, 200) +
  ggtitle("Proportion part = 7\nfruit with insect emergence hole") -> p2

p1 + p2 + plot_annotation(
  title = "Thinouia myriantha")

```

The distributions look kind of similar, can we measure this quantitatively?

Kolmogorov–Smirnov test: What is the probability that these two sets of samples were drawn from the same (but unknown) probability distribution?

```{r}

trap_dat_caps %>%
  ungroup() %>%
  filter(sp4 == "thim") %>% 
  filter(total_seeds >=5) %>%
  select(proportion_abscised) -> x

thim_part7 %>% 
  ungroup() %>%
  filter(total_seeds >=5) %>%
  select(proportion_insect_abscised) -> y

ks.test(x$proportion_abscised, y$proportion_insect_abscised)

```

Null hypothesis: x and y were drawn from the same continuous distribution. We have p < 0.05 which means we reject the null hypothesis - the two samples came from different distributions.

We can also use a Mann-Whitney test (equivalent to the Wilcoxon rank sum test), which is better at handling tied values.

```{r}
wilcox.test(x$proportion_abscised, y$proportion_insect_abscised, conf.int = TRUE, paired = FALSE)
```

Here, the null hypothesis is that the distributions of x and y differ by a location shift of mu and the alternative is that they differ by some other location shift. We have p < 0.05 which means we reject the null hypothesis - the two samples came from different distributions.

__To do:__ Repeat for species with highest counts of part 7, top 10?
