---
title: "Changing edge effect mitigation method"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  github_document:
    keep_html: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/13_changing_edge_effects/")
```

Since we have changed alpha, we'll probably need to change our radius for correcting for edge effects too. I'll re-run the exercise I did here [03_edge-effects.md](03_edge-effects.md).

```{r, warning=FALSE, message=FALSE}
library("tidyverse"); theme_set(theme_bw(base_size=10))
library("sf")
library("rdist")
library("knitr")
library("patchwork")
library("here")

trap_connect <- readRDS(here::here("data", "clean", "trap_connect_30m.rds"))

tree_data <- readRDS(here::here("data", "clean", "tree_data.rds"))

```


```{r random-sampling, warning=FALSE}
## TRAP DATA

# subset to traps which are on the LHS of the plot 
# and not near the 'real' edges
centre_traps <- subset(trap_connect, x < 500 & x > 100 & y < 400 & y > 100)

# randomly select 1000 of these traps
sample_n(centre_traps, 4000) %>% 
  rename(CI_truth = connectivity) -> test_traps

## TREE DATA

# subset tree data to only include trees on the LHS of the plot
subset(tree_data, x < 500) -> tree_data_50

```

```{r observed-connectivity, warning=FALSE, message=FALSE}
## First calculate eulcidian distances

calculate_dist <- function (species, yr) {
	# subset tree data
	bd <- dplyr::filter(tree_data_50,
	                    (tree_data_50[,"sp4"] == species) & (tree_data_50[,"year"] == yr) )

	# subset trap data
	td <- dplyr::filter(test_traps,
	                    (test_traps[,"sp4"] == species) & (test_traps[,"year"] == yr) )

	# create distance matrix using x and y co-ordinates
	dists <- cdist(bd[,c("x", "y")], td[,c("x", "y")], metric = "euclidean")
	dists_df <- as.data.frame(dists)

	# label col names with trap IDs
	colnames(dists_df) <- unlist(td$trap)

	# bind with the tree data
	cbind(bd, dists_df)
}

# create lists of sp and year to pass to function
test_traps %>%
  distinct(year) %>%
  inner_join(distinct(tree_data, year)) %>%
  pull(year) -> year_list

test_traps %>%
  distinct(sp4) %>%
  inner_join(distinct(tree_data, sp4)) %>%
  pull(sp4) -> sp_list

# apply function to all pairwise combinations of year and species
# will return list of dfs
all_dists <- outer(sp_list, year_list, FUN = Vectorize(calculate_dist))

bci_dists <- dplyr::bind_rows(all_dists)

## Now calculate Hanski's connectivity index

calculate_CI <- function (trap) {
	drop_na(bci_dists, trap) %>%
	group_by(tree, year) %>% 
	mutate(a = exp( (-1/30) * eval(parse(text=trap)) ) * dbh^(0.5)) %>% 
	ungroup() %>%
	dplyr::select(year, sp4, trap, a) %>%
	group_by(year, sp4) %>%
	summarise(trap = paste(trap), CI_obs = sum(a), .groups = "drop")
}

test_traps %>%
  distinct(trap) %>%
  pull(trap) -> trap_list

CI_data <- parallel::mclapply(trap_list, calculate_CI, mc.cores = 3)

CI_data_b <- bind_rows(CI_data)

# join our values of observed CI back to our values of true CI
left_join(CI_data_b, test_traps, by = c("year", "sp4", "trap")) -> sample

```

## Predict connectivity

```{r predict-connectivity, warning=FALSE}
correct_for_edge <- function(species, yr, radius){
	# draw the fdp
	fdplot <- rbind(c(0,0), c(0, 500), c(500, 500), c(500, 0), c(0,0))
	# make it a polygon
	fdplot_sf <- sf::st_polygon(list(fdplot))
	# subset edgeTraps to sp i and yr i
	dplyr::filter(sample, 
		(.data[["sp4"]] == species) & (.data[["year"]] == yr)) %>%
	# make traps sf points
	sf::st_as_sf(coords = c("x", "y")) %>% 
	# create a radius around the traps
	mutate(buf = st_buffer(geometry, dist = radius)) %>% 
	# calculate the area of intersection between each radius & the fdp
	mutate(area = sf::st_area( 
		sf::st_intersection(.data[["buf"]], fdplot_sf) )) %>% 
	# subset to traps that have an intersection / are on the edge
	filter(area > 0 & area < ( (pi * radius^2) - 1) ) %>%
	# calculate our prediction
	mutate(CI_pred = (CI_obs/area) * (pi * radius^2), radius = radius )
}

# create df of every combination of year, sp and radius
sample %>% 
	count(year, sp4) %>% 
	select(sp4, year) %>% 
	rename(species = sp4, yr = year)	-> sample_keys

radius_keys <- tibble(radius = c(30, 60, 90, 120, 150, 180))

sample_radius_keys <- merge(sample_keys, radius_keys)

# pass to function
test_traps_adjusted <- pmap(sample_radius_keys, correct_for_edge)
```

## Compare radii

```{r plot, warning=FALSE, fig.width=10, fig.height=10}
# bind output dfs and calculate error
bind_rows(test_traps_adjusted) %>%
	mutate(CI_error = CI_pred - CI_truth, radius = as.factor(radius)) -> output

# add a 'control' 
sample %>%
	mutate(CI_error = CI_obs - CI_truth, radius = as.factor("control")) %>%
	select(year, sp4, trap, CI_error, radius) -> control

bind_rows(output, control) -> plotting_data

# plot error as jitter
ggplot(plotting_data) +
  geom_jitter(aes(x = radius, y = CI_error)) +
	labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p1

# as violins
ggplot(plotting_data) +
  geom_violin(aes(x = radius, y = CI_error)) +
  labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p2

# with x-axis ordered by decreasing median error
ggplot(plotting_data) +
  geom_jitter(aes(x = reorder(radius, abs(CI_error), median), y = CI_error)) +
  labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p3

ggplot(plotting_data) +
  geom_violin(aes(x = reorder(radius, abs(CI_error), median), y = CI_error)) +
  labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p4

(p1 + p2) / (p3 + p4)
```

As the radius decreases we are getting fewer points because a smaller radius means that the radius is less likely to intersect with the edge of the plot. We're going to have to create more artificial edges, perhaps sweep across in 50m increments. 

Or maybe we could just use a bigger sample size? We are currently just using 1000 trap-sp-years.. we have many more to play with and the edge-sweep idea is proving difficult to pull off. I might need to run it on the cluster though.

```{r warning=FALSE}
# not near the 'real' edges and only survey years
trap_connect %>%
  filter(year == 1990 | year == 1995 |
           year == 2000 |year == 2005 |
           year == 2010 |year == 2015) %>%
  filter(x < 900 & x > 100 & y < 400 & y > 100) -> centre_traps_2

# randomly select 1000 traps
sample_n(centre_traps_2, 1000) %>% 
 rename(CI_truth = connectivity) -> test_traps_2

## First calculate eulcidian distances

calculate_dist_2 <- function (species, yr) {
	# subset tree data
	bd <- dplyr::filter(tree_data,
	                    (tree_data[,"sp4"] == species) & (tree_data[,"year"] == yr) )

	# subset trap data
	td <- dplyr::filter(test_traps_2,
	                    (test_traps_2[,"sp4"] == species) & (test_traps_2[,"year"] == yr) )

	# create distance matrix using x and y co-ordinates
	dists <- cdist(bd[,c("x", "y")], td[,c("x", "y")], metric = "euclidean")
	dists_df <- as.data.frame(dists)

	# label col names with trap IDs
	colnames(dists_df) <- unlist(td$trap)

	# bind with the tree data
	cbind(bd, dists_df)
}

# create lists of sp and year to pass to function
test_traps_2 %>%
  distinct(year) %>%
  inner_join(distinct(tree_data, year)) %>%
  pull(year) -> year_list_2

test_traps_2 %>%
  distinct(sp4) %>%
  inner_join(distinct(tree_data, sp4)) %>%
  pull(sp4) -> sp_list_2

# apply function to all pairwise combinations of year and species
# will return list of dfs
all_dists_2 <- outer(sp_list_2, year_list_2, FUN = Vectorize(calculate_dist_2))

bci_dists_2 <- dplyr::bind_rows(all_dists_2)

## Now calculate observed connectivity with different edges

calculate_CI <- function (trap, edge) {
  filter(bci_dists_2, x < edge) %>%
	drop_na(trap) %>%
	group_by(tree, year) %>% 
	mutate(a = dbh * exp( (-1/30) * eval(parse(text=trap)) ) ) %>% 
	ungroup() %>%
	dplyr::select(year, sp4, trap, a) %>%
	group_by(year, sp4) %>%
	summarise(trap = paste(trap), CI_obs = sum(a), edge = edge, .groups = "drop")
}

test_traps_2 %>%
  distinct(trap) %>%
  pull(trap) -> trap_list_2

edge_keys_2 <- tibble(edge = c(800, 700, 600, 500, 400, 300, 200))

merge(trap_list_2, edge_keys_2) %>%
  rename(trap = x) -> keys_trap_edge_2

CI_data_2 <- pmap(keys_trap_edge_2, calculate_CI)

CI_data_b_2 <- bind_rows(CI_data_2)

# join our values of observed CI back to our values of true CI
left_join(CI_data_b_2, test_traps_2, by = c("year", "sp4", "trap")) %>%
  filter(CI_obs != CI_truth) -> sample_2

split_plot_by_edge <- function(species, yr, edge) {
  fdplot <-
    rbind(c(100, 100), c(100, 400), c(edge, 400), c(edge, 100), c(100, 100))
  
  fdplot_sf <- sf::st_polygon(list(fdplot))
  
  sample_2 %>% 
    filter(x < edge) %>%
    filter((.data[["sp4"]] == species) & (.data[["year"]] == yr) &
             (.data[["edge"]] == edge)) %>%
    mutate(plot = list(fdplot))
}

correct_for_edge <- function(data, radius) {
  data %>%
    sf::st_as_sf(coords = c("x", "y")) %>%
    rowwise() %>%
    mutate(buf = sf::st_buffer(geometry, dist = radius)) %>%
    mutate(intersection = sf::st_intersects(buf, sf::st_polygon(list(plot)))) %>%
    mutate(intersection = list(length(unlist(intersection)) != 0)) %>%
    filter(intersection == TRUE) %>%
    mutate(area = sf::st_area(sf::st_intersection(.data[["buf"]],
                                                  sf::st_polygon(list(
                                                    plot
                                                  ))))) %>%
    ungroup() %>%
    filter(area >= 0 & area < ((pi * radius ^ 2) )) %>%
    rowwise() %>%
    mutate(CI_pred = (CI_obs / area) * (pi * radius ^ 2),
           radius = radius) %>%
    ungroup()
}
  
# create df of every combination of year sp and edge
sample_2 %>% 
	count(year, sp4, edge) %>%
	select(sp4, year, edge) %>% 
	rename(species = sp4, yr = year) -> keys_species_yr_edge

# pass to function
test_traps_adjusted_2 <- pmap(keys_species_yr_edge, split_plot_by_edge)

test_traps_adjusted_2_bind <- bind_rows(test_traps_adjusted_2)
keys_radius_2 <- tibble(radius = c(10, 20, 30, 40, 50))

# pass to function
test_traps_done <- pmap(keys_radius_2, 
                              correct_for_edge, data = test_traps_adjusted_2_bind)

bind_rows(test_traps_done) %>%
	mutate(CI_error = CI_pred - CI_truth, radius = as.factor(radius)) -> output

sample_2 %>%
	mutate(CI_error = CI_obs - CI_truth, radius = as.factor("control")) %>%
	select(year, sp4, trap, CI_error, radius) -> control

bind_rows(output, control) -> plotting_data

```

```{r plot-2, warning=FALSE, fig.width=10, fig.height=10}

# plot error as jitter
ggplot(plotting_data) +
  geom_jitter(aes(x = radius, y = CI_error)) +
	labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p1

# as violins
ggplot(plotting_data) +
  geom_violin(aes(x = radius, y = CI_error)) +
  labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p2

# with x-axis ordered by decreasing median error
ggplot(plotting_data) +
  geom_jitter(aes(x = reorder(radius, abs(CI_error), median), y = CI_error)) +
  labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p3

ggplot(plotting_data) +
  geom_violin(aes(x = reorder(radius, abs(CI_error), median), y = CI_error)) +
  labs(x = "radius", y = "error") +
  geom_hline(yintercept  = 0, colour = "red", linetype = 2) -> p4

(p1 + p2) / (p3 + p4)

```

