---
title: "Compare models"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  github_document:
    keep_html: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/26_model-comparison/")
```

Compare models with all 3 connectivity/density measures.
Do we actually need zero inflation?

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("brms")
library("patchwork")
```

```{r}
zibb <-
  readRDS(here::here("output", "models", "pheno-repro-adjust",
                     "full_conn_zib.rds"))

bbinom <-
  readRDS(here::here("output", "models", "pheno-repro-adjust",
                     "full_conn_binom.rds"))

zoib <-
  readRDS(here::here("output", "models", "pheno-repro-adjust",
                     "full_conn_simple.rds"))
```

## `summary()`

```{r}
summary(zibb)
```

```{r}
summary(bbinom)
```

```{r}
summary(zoib)
```


## Do we need zero inflation?

```{r}
pp_check(zibb, type = "stat", stat = function(y) mean(y == 0), ndraws = 500) +
  pp_check(bbinom, type = "stat", stat = function(y) mean(y == 0), ndraws = 500) +
  patchwork::plot_layout(guides = "collect")
  
```

Looks like the ZIBB didn't fit very well, 
but the BBinom is slightly under predicting zeros.

```{r}
# draw posterior predictive samples (matrix draws x observations)
pp_bbinom <- posterior_predict(bbinom, ndraws = 2000)  

# observed zeros
obs_zero <- as.integer(bbinom$data$abscised_seeds == 0)

# compute predicted overall zero proportion per draw
pred_zero_rate <- apply(pp_bbinom, 1, function(draw) mean(draw == 0))

# summary
obs_zero_rate <- mean(obs_zero)
median_pred <- median(pred_zero_rate)
ci_pred <- quantile(pred_zero_rate, c(0.025, 0.975))

tibble(
  observed = obs_zero_rate,
  pred_median = median_pred,
  pred_lo = ci_pred[1],
  pred_hi = ci_pred[2]
)
```

If observed value is within 95% CI, the model can predict the observed zero rate.

Check zeros by `total_seeds`:

```{r}
# choose bins (quantile binning)
df2 <- bbinom$data %>%
  mutate(n_bin = ntile(total_seeds, 50)) %>%   # 5 bins; change as needed
  group_by(n_bin) %>%
  mutate(mean_n = mean(total_seeds)) %>%
  ungroup()

# map original indices to bins
bins <- df2$n_bin

# compute predicted zero rate per bin for each draw
pred_zero_by_bin <- apply(pp_bbinom, 1, function(draw) {
  tapply(draw == 0, bins, mean)
}) # result: draws x bins (list->matrix)

# summarize per bin
bin_summary <- as_tibble(t(pred_zero_by_bin)) %>%
  mutate(draw = row_number()) %>%
  pivot_longer(-draw, names_to = "n_bin", values_to = "pred_zero") %>%
  group_by(n_bin) %>%
  summarise(
    pred_median = median(pred_zero),
    pred_lo = quantile(pred_zero, 0.025),
    pred_hi = quantile(pred_zero, 0.975)
  ) %>%
  mutate(n_bin = as.integer(n_bin)) %>% 
  arrange(n_bin) 

# observed by bin
obs_by_bin <- df2 %>%
  group_by(n_bin, mean_n) %>%
  summarise(obs_zero = mean(abscised_seeds == 0), .groups = "drop") %>%
  arrange(as.integer(n_bin))

# join and plot
check_bin <- left_join(obs_by_bin, bin_summary, by = "n_bin")

ggplot(check_bin, aes(x = mean_n)) +
  geom_ribbon(aes(ymin = pred_lo, ymax = pred_hi), alpha = 0.2) +
  geom_line(aes(y = pred_median), linetype = "dashed") +
  geom_point(aes(y = obs_zero), size = 2) +
  labs(x = "mean total_seeds in bin", y = "proportion zeros",
       title = "Observed vs predicted zero proportion by total_seeds bin") +
  theme_minimal()
```

For very large `total_seeds` the beta-binomial predicts more zeros than actually observed. This suggests the zeros are not structural. 
If they were, we'd see too many zeros overall and *underprediction* of zeros 
at large n.

This pattern suggests for large `total_seeds` observing zero abscised seeds 
is very unlikely.

-> trees which are very productive likely have a larger proportion of abscised seeds 
