---
title: "Testing different values of alpha"
author: "Eleanor Jackson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    keep_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "figures/28_alpha-radii/")
ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r message=FALSE, warning=FALSE}
library("tidyverse"); theme_set(theme_bw(base_size = 10))
library("sf")
library("patchwork")
```

Use an effective radius that captures almost all kernel weight 
(so anything beyond it contributes negligible connectivity).

effective radius (capturing 95% of influence) = -ln(0.05) / alpha

```{r}
test_data <- 
  readRDS(here::here("data", "clean", "connect_all_alpha.rds"))
```

```{r}

eps <- 0.05 #captures 95% of influence

# draw the fdp
fdplot <- rbind(c(0,0), c(0, 500), c(1000, 500), c(1000, 0), c(0,0))

# make it a polygon
plot_polygon <- sf::st_polygon(list(fdplot))

plot(plot_polygon)
```

```{r}
dat_sf <- sf::st_as_sf(test_data, coords = c("x", "y"),
                       remove = FALSE) 

# distance to boundary
dist_to_boundary <- st_distance(dat_sf, st_boundary(plot_polygon),
                                which = "Euclidean")

# st_distance returns units; convert to numeric meters
dist_to_boundary_m <- as.numeric(dist_to_boundary)

dat_sf$dist_to_edge_m <- dist_to_boundary_m

# effective radii (exponential kernel)
# the distance where the weight drops to some small threshold

dat_sf <- 
  dat_sf %>% 
  mutate(r_eff =  (- log(eps)) / alpha) %>% 
  mutate(location = ifelse(dist_to_edge_m >= r_eff, "interior",
                       "edge"))

dat_sf %>% 
  sf::st_drop_geometry() %>% 
  mutate(dist = 1/alpha) %>% 
  select(trap, x, y , location, alpha, dist) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = location)) +
  geom_point() +
  facet_wrap(~dist)

```

```{r}
dat_sf %>% 
  mutate(alpha = 1/alpha) %>%
  ggplot(aes(x = conn_RH, colour = location)) +
  geom_density() +
  facet_wrap(~alpha, scales = "free") +
  ggtitle("Reproductive heterospecifics")

```

Traps at the edge generally have connectivity distributions which
are shifted lower. 
This is more apparent when the "average migration distance" (or 1/alpha) is low,
because there are fewer "edge traps" in these datasets and they are closer
to the edge (see previous figure).

```{r}
dat_sf %>% 
  sf::st_drop_geometry() %>% 
  filter(alpha == 1/40 & sp4 == "jacc") %>% 
  filter(year == 1990 | year == 2000 | year == 2010| year == 2020 ) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = conn_RH)) +
  geom_point() +
  facet_wrap(~year) +
  ggtitle("jacc")

dat_sf %>% 
  sf::st_drop_geometry() %>% 
  filter(alpha == 1/40 & sp4 == "alsb") %>% 
  filter(year == 1990 | year == 2000 | year == 2010| year == 2020 ) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = conn_RH)) +
  geom_point() +
  facet_wrap(~year) +
  ggtitle("alsb")

dat_sf %>% 
  sf::st_drop_geometry() %>% 
  filter(alpha == 1/40 & sp4 == "hybp") %>% 
  filter(year == 1990 | year == 2000 | year == 2010| year == 2020 ) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = conn_RH)) +
  geom_point() +
  facet_wrap(~year) +
  ggtitle("hybp")
```

Traps on the RHS of the plot seem to have a high density of heterospecifics.
Perhaps just many trees in this area - near the river?

This is causing the bimodal distribution in `conn_RH` values at higher
alpha.

```{r}
dat_sf %>% 
  mutate(alpha = 1/alpha) %>%
  ggplot(aes(x = conn_RC, colour = location)) +
  geom_density() +
  facet_wrap(~alpha, scales = "free") +
  ggtitle("Reproductive conspecifics")

```

Edge traps also have lower reproductive conspecific density,

```{r}
dat_sf %>% 
  mutate(alpha = 1/alpha) %>%
  ggplot(aes(x = conn_NRC, colour = location)) +
  geom_density() +
  facet_wrap(~alpha, scales = "free") +
  ggtitle("Non-reproductive conspecifics")

```

and lower non-reproductive conspecific density.

```{r}
# dat_sf %>%
#   sf::st_drop_geometry() %>%
#   mutate(location = as.factor(location)) %>%
#   saveRDS(here::here("data", "clean", "connect_all_alpha_buffer.rds"))
```

How different is the new radii method from excluding those which are less than
1/alpha from the edge?

```{r}
dat_sf %>% 
  sf::st_drop_geometry() %>% 
  mutate(dist = 1/alpha) %>% 
  filter(dist == 20) %>% 
  select(trap, x, y , location, alpha, dist) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = location)) +
  geom_point() +
  
  dat_sf %>% 
  filter(alpha == 1/20) %>%
  select(trap, x, y , location, alpha) %>% 
  mutate(location = case_when(
    x < 980 & x > 20 & y < 480 & y > 20 ~ "interior",
    .default = "edge"
  )) %>% 
  mutate(location = as.factor(location)) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = location)) +
  geom_point() +
  
  plot_layout(ncol = 1)

```

