---
title: "Testing different values of alpha"
author: "Eleanor Jackson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    keep_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path = "figures/28_alpha-radii/")
ggplot2::theme_set(ggplot2::theme_classic(base_size = 10))
```

```{r message=FALSE, warning=FALSE}
library("tidyverse"); theme_set(theme_bw(base_size = 10))
library("sf")
library("patchwork")
```

Use an effective radius that captures almost all kernel weight 
(so anything beyond it contributes negligible connectivity).

effective radius (capturing 95% of influence) = -ln(0.05) / alpha

```{r}
test_data <- 
  readRDS(here::here("data", "clean", "connect_all_alpha.rds"))
```

```{r}

eps <- 0.05 #captures 95% of influence

# draw the fdp
fdplot <- rbind(c(0,0), c(0, 500), c(1000, 500), c(1000, 0), c(0,0))

# make it a polygon
plot_polygon <- sf::st_polygon(list(fdplot))

plot(plot_polygon)
```

```{r}
dat_sf <- sf::st_as_sf(test_data, coords = c("x", "y"),
                       remove = FALSE) 

# distance to boundary
dist_to_boundary <- st_distance(dat_sf, st_boundary(plot_polygon),
                                which = "Euclidean")

# st_distance returns units; convert to numeric meters
dist_to_boundary_m <- as.numeric(dist_to_boundary)

dat_sf$dist_to_edge_m <- dist_to_boundary_m

# effective radii (exponential kernel)
# the distance where the weight drops to some small threshold

dat_sf <- 
  dat_sf %>% 
  mutate(r_eff =  (- log(eps)) / alpha) %>% 
  mutate(location = ifelse(dist_to_edge_m >= r_eff, "interior",
                       "edge"))

dat_sf %>% 
  sf::st_drop_geometry() %>% 
  mutate(r_eff_m = round(r_eff)) %>% 
  select(trap, x, y , location, alpha, r_eff_m) %>% 
  distinct() %>% 
  ggplot(aes(x = x, y = y, colour = location)) +
  geom_point() +
  facet_wrap(~r_eff_m)

```

```{r}
dat_sf %>% 
  mutate(r_eff_m = round(r_eff)) %>% 
  ggplot(aes(x = conn_RH, colour = location)) +
  geom_density() +
  facet_wrap(~r_eff_m, scales = "free") +
  ggtitle("Reproductive heterospecifics")

```

Traps at the edge generally have connectivity distributions which
are shifted lower. 

```{r}
dat_sf %>% 
  mutate(r_eff_m = round(r_eff)) %>% 
  ggplot(aes(x = conn_RC, colour = location)) +
  geom_density() +
  facet_wrap(~r_eff_m, scales = "free") +
  ggtitle("Reproductive conspecifics")

```

Edge traps also have lower reproductive conspecific density,

```{r}
dat_sf %>% 
  mutate(r_eff_m = round(r_eff)) %>% 
  ggplot(aes(x = conn_NRC, colour = location)) +
  geom_density() +
  facet_wrap(~r_eff_m, scales = "free") +
  ggtitle("Non-reproductive conspecifics")

```

and lower non-reproductive conspecific density.

```{r}
# dat_sf %>%
#   sf::st_drop_geometry() %>%
#   mutate(location = as.factor(location)) %>%
#   saveRDS(here::here("data", "clean", "connect_all_alpha_buffer.rds"))
```
