---
title: "Insect rearing data"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: 
  github_document:
    keep_html: false
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.path="figures/37_insect-rearing-data/")
```

```{r, warning=FALSE, message=FALSE}
library("tidyverse")
library("patchwork")
library("glmmTMB")
```

Reviewers are concerned that the causality between immature seed mortality and natural
enemies is not established.

Sofia: We could perhaps tap into the data collected in the context of [our food web study](https://onlinelibrary.wiley.com/doi/abs/10.1111/ele.13359). We have insect rearing data and fruit/seed dissection data which is assembled separately for mature and immature fruits. The data sets are not perfect, but should still allow us to demonstrate that for most species, immature fruits are *way* more likely to yield insects (or show signs of insect attack) than mature fruits, which could perhaps help us in building the case.
 
1. Seeds_saved_for_rearing(…).xlsx: Contains a column called Nr_units_predated which shows data from dissections (rather conservative – sometimes the material was in quite a bad state after 3 months of rearing, so hard to see evidence of predation). Metadata worksheet provides column descriptions.

2. Insects_final_cleaned(…).xlsx: Each row shows the number of insects of a given morphospecies collected from a given pot on a given day (and therefore placed in the same vial for storage) – if lots of insects there might be multiple vials (and rows in the file) for a morphospecies x rearing pot x day combination. The column Quantity gives the number of individuals. SeedPot column is the same as Pot_ID in the previous file. (You will need to link the two to get information on whether seeds/fruits were mature or immature on collection.

3. Seed predator summary(…).xlsx: I can’t remember why I did this summary and hope that it does not contain any errors, but it looks like some info here might be potentially useful. Worksheet IndsPerMorphospPerMaturityStage seems potentially useful, and IndsPerMorphosp_seedpreds only has filtered out morphospecies that we believe are feeding on seeds (not the fleshy parts of the fruit like e.g. most flies and some moth families do). Could possibly remove morphospecies that don’t appear in this worksheet from data summaries?

With this data, we can ask if more insects are reared from immature than from mature fruits/seeds.
Can also restrict insects to seed predators and exclude fruit-feeding insects.
Our response is: insect emergence rate from pots containing immature versus mature fruits/seeds 
i.e. number of emerged insects divided by number of seeds.

Read in the data:

```{r, warning=FALSE}
data_seed <- readxl::read_xlsx(
  here::here("data", "raw", "insect-data", 
             "Seeds_saved_for_rearing_Cleaned_June2014.xlsx"),
  na = c("", "NA")) |> 
  janitor::clean_names()

data_insect <- readxl::read_xlsx(
  here::here("data", "raw", "insect-data", 
             "Insects_final_cleaned_September2015.xlsx"),
  na = c("", "NA")) |> 
  janitor::clean_names()

seed_pred_morphos <- readxl::read_xlsx(
  sheet = 4,
  here::here("data", "raw", "insect-data", 
             "Seed_predator_summary_16July2014.xlsx")) |> 
  janitor::clean_names() |> 
  filter(seed_predator == "Yes")
```

It would be good to estimate the number of seeds within collected fruits.

```{r}
# Joe's fruit dissection data
seed_trait <-
  read_tsv(
    here::here(
      "data",
      "raw",
      "seed-masses",
      "Fruit_Seed_masses_20210111_BanisteriopsisCorrectedToBactrismajor.txt"
    )
  ) %>%
  rename_with(tolower) %>%
  mutate(id = paste(sp4, ind.unique, fruit.unique, sep = "_")) %>%
  select(sp4, id, n_seedfull, n_capsules) %>%
  distinct() %>%
  group_by(sp4) %>%
  summarise(seeds_per_fruit = mean(n_seedfull, na.rm = TRUE)) |> 
  select(sp4, seeds_per_fruit)

```

Calculate seeds per pot:

```{r}
data_seed <- 
  data_seed |> 
  left_join(seed_trait, by = c("codigo" = "sp4")) |> 
  rowwise() |> 
  mutate(nr_equ_seeds = 
           ifelse(unit == "Fruit", nr_fruits * seeds_per_fruit, nr_seeds)) 
```

There 124 species out of 483 for which we do not have data on "seeds per fruit".

```{r}
# join pot data with insect data
insects_by_pot <- 
  data_insect %>%
  filter(!is.na(seed_pot)) %>%
  group_by(seed_pot) %>%
  summarise(
    n_insects = sum(quantity, na.rm = TRUE),
    .groups = "drop"
  )

data_join <-
  data_seed |> 
  left_join(insects_by_pot, by = join_by("pot_id" == "seed_pot")) |> 
  mutate(n_insects = replace_na(n_insects, 0)) |> 
  filter(nr_equ_seeds > 0) 

data_join |> 
  ggplot(aes(x = n_insects, fill = maturity_type)) +
  geom_density(alpha = 0.4) +
  ylab("N pots") +
  coord_cartesian(xlim = c(0,20))
```

Most pots had zero insects.

```{r}
data_join |> 
  group_by(maturity_type) |> 
  summarise(sum_seeds = sum(nr_equ_seeds, na.rm = TRUE),
            n_insects = sum(n_insects, na.rm = TRUE),
            sum_fruits = sum(nr_fruits, na.rm = TRUE)) |> 
  mutate(seed_emergence_rate = n_insects/sum_seeds,
         fruit_emergence_rate = n_insects/sum_fruits) |> 
  print(width = Inf)
```

Emergence rate is higher for mature material, 
this would mean 0.02 insects emerged per mature seed on average vs
0.007 per immature seed and
0.33 per mature fruit vs 0.09 per immature fruit.

Try restricting to only seed predators:

```{r}
insects_by_pot_seedpreds <- 
  data_insect %>%
  filter(morphospecies %in% seed_pred_morphos$morphospecies) |> 
  filter(!is.na(seed_pot)) %>%
  group_by(seed_pot) %>%
  summarise(
    n_insects = sum(quantity, na.rm = TRUE),
    .groups = "drop"
  ) 

data_join_seedpreds <-
  data_seed |> 
  left_join(insects_by_pot_seedpreds, by = join_by("pot_id" == "seed_pot")) |> 
  mutate(n_insects = replace_na(n_insects, 0))|> 
  filter(nr_equ_seeds > 0)

data_join_seedpreds |> 
  group_by(maturity_type) |> 
  summarise(sum_seeds = sum(nr_equ_seeds, na.rm = TRUE),
            n_insects = sum(n_insects, na.rm = TRUE),
            sum_fruits = sum(nr_fruits, na.rm = TRUE)) |> 
  mutate(seed_emergence_rate = n_insects/sum_seeds,
         fruit_emergence_rate = n_insects/sum_fruits) |> 
  print(width = Inf)
```

Try fitting a model to account for differences in sampling effort and species.

First for fruits only:

```{r}
data_fit <- data_join_seedpreds %>%
  mutate(
    maturity_type = factor(maturity_type),
    maturity_type = relevel(maturity_type, ref = "maduro")
  )

mfruit <- glmmTMB(
  n_insects ~ maturity_type +
    offset(log(nr_fruits)) +
    (1 | codigo),
  family = nbinom2,
  data = filter(data_fit, nr_fruits > 0)
)

summary(mfruit)
exp(confint(mfruit, parm = "maturity_typeinmaduro"))
```

The fitted insect emergence rate per fruit in immature material is about 52% lower than in mature fruits.

Now including seeds plus the estimated seeds within fruits (nr_equ_seeds):

```{r}
m1 <- glmmTMB(
  n_insects ~ maturity_type +
    offset(log(nr_equ_seeds)) +
    (1 | codigo),
  family = nbinom2,
  data = drop_na(data_fit, seeds_per_fruit)
)

summary(m1)
exp(confint(m1, parm = "maturity_typeinmaduro"))
```

The estimated insect emergence rate for immature material is 0.58 times the rate for mature material. i.e., the emergence rate is about 42% lower in immature material, and it's a significant effect

```{r}
data_seed %>%
  filter(nr_equ_seeds > 0) |> # only pots containing seeds
  group_by(codigo) %>%
  summarise(
    n_maturity = n_distinct(maturity_type),
    immature = any(maturity_type == "inmaduro"),
    mature   = any(maturity_type == "maduro"),
    .groups = "drop"
  ) %>%
  count(n_maturity)
```

Not all species contain both maturity types.
If species and `maturity_type` are largely confounded,
much of the difference between immature and mature material 
is probably associated with which plant species occur in each maturity category, 
rather than maturity alone.

We could try restricting the analysis to only the species which have both immature
and mature fruits.

```{r}
species_both <- 
  data_fit %>%
  group_by(codigo) %>%
  filter(n_distinct(maturity_type) == 2) %>%
  ungroup()
```


```{r}
m_both <- glmmTMB(
  n_insects ~ maturity_type +
    offset(log(nr_equ_seeds)) +
    (1 | codigo),
  family = nbinom2,
  data =  drop_na(species_both, seeds_per_fruit)
)

summary(m_both)
exp(fixef(m_both)$cond)
exp(confint(m_both, parm = "maturity_typeinmaduro"))
```

This is similar to the previous model result: the estimated insect emergence rate for immature material is 0.58 times the rate for mature material.

What about seeds which show signs of insect attack/ emergence holes?

The metadata say:
"Nr_units_predated:	Number of units (seeds or fruits) with clear signs of predation (scored when sample discarded). Unless the following column says otherwise - all units (seeds or fruits) were examined."


```{r}
data_seed_attack <- 
  data_seed |> 
  drop_na(seeds_per_fruit) |> 
  # convert n examined fruit to est seeds
  mutate(nr_eq_seeds_examined_for_predation = 
           ifelse(unit == "Fruit", 
                  nr_units_examined_for_predation * seeds_per_fruit,
                  nr_units_examined_for_predation)) |> 
  # if NA, all units were examined
  mutate(nr_eq_seeds_examined_for_predation = 
           coalesce(
             nr_eq_seeds_examined_for_predation, 
             nr_equ_seeds)
         ) |>
  # convert n predated fruit to est seeds
  mutate(nr_eq_seeds_predated = 
           ifelse(
             unit == "Fruit", 
             nr_units_predated * seeds_per_fruit, 
             nr_units_predated)
         ) |> 
  filter(nr_eq_seeds_predated <= nr_eq_seeds_examined_for_predation) |> 
  mutate(
    nr_eq_seeds_examined_for_predation = round(nr_eq_seeds_examined_for_predation),
    nr_eq_seeds_predated = round(nr_eq_seeds_predated)) |> 
  filter(nr_eq_seeds_examined_for_predation > 0) |> 
  mutate(
    maturity_type = factor(maturity_type),
    maturity_type = relevel(maturity_type, ref = "maduro")
  ) |> 
  drop_na(nr_eq_seeds_predated)

data_seed_attack |> 
  group_by(maturity_type) |> 
  summarise(n_predated = sum(nr_eq_seeds_predated, na.rm = TRUE),
            sum_seeds = sum(nr_eq_seeds_examined_for_predation, na.rm = TRUE)) |> 
  mutate(proportion_predated = (n_predated/sum_seeds)*100) 

```

```{r}
m_attack <- glmmTMB(
	cbind(
		nr_eq_seeds_predated,
		nr_eq_seeds_examined_for_predation - nr_eq_seeds_predated
	) ~ maturity_type +
		(1 | codigo),
	family = betabinomial(link = "logit"),
	data = data_seed_attack
)

m_attack
exp(confint(m_attack, parm = "maturity_typeinmaduro"))
```

The model estimates that the odds of a seed showing signs of predation in immature material are about 17% lower than in mature material -- the opposite of what we would expect -- but a large confidence interval.
