---
title: "Connectivity"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: lumen
    highlight: pygments
---

I'm going to try two different measures of connectivity, __Hanski’s Connectivity Index (CI)__ (Hanski 1994) and the __Conspecific neighbourhood fecundity index (NFI)__ (Jones & Comita 2008). 

They differ in that Hanski's CI includes an exponential function, meaning that it doesn't weigh short distances as heavily. (e.g if a tree was right next to a trap you'd risk getting an inf value for NFI). CI also needs the average dispersal distance to scale your distances, which I'm just going to take as 100m for now.

I had to alter these equations a little. NFI uses the yearly mean total fruit set per tree to weight individuals - we don't have this, so I just used DBH, which assumes that larger individuals produce more fruit.

__Hanksi's Connectivity Index__\

$CI = \sum exp(-\, \alpha \, dist_{ji}) A_j$ \

__Conspecific neighbourhood fecundity index__\

$NFI = \sum \frac {A_j}{dist_{ij}}$ \

_j_ = conspecific tree within a 100m radius of the trap\
_dist_ = distance between trap i and tree j\
$\alpha$ = 1/ avg. dispersal ability\
_A_ = DBH of tree j\

```{r, warning=FALSE, message=FALSE}
rm(list = ls())

require(tidyverse)
library(knitr)
library(rdist)
opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy=FALSE, cache.lazy = FALSE,cache = TRUE)
```
## Read in and compile the data
We are going to focus on Cordia bicolour in 2015 to begin with.

```{r, warning=FALSE}
load("../data/bci.tree/bci.tree8.rdata")
load("../data/bci.spptable.rdata")
trapDat <- read.csv("../output/tables/proportionAbscisedPerTrap.csv") 
RDA <- read.csv("../output/tables/R50.csv") 

# sort out date formating and create a year column
bci.tree8$date <- as.Date(bci.tree8$date, 
	origin="1960-01-01", format= "%Y-%m-%d")
bci.tree8$year <- as.integer(format(as.Date(bci.tree8$date), "%Y"))

bci8 <- left_join(bci.tree8, bci.spptable, by = "sp")
bci8 <- left_join(bci8, RDA, by = "sp")

# add id tags to id values so they don't just look like numbers
trapDat$trap <- paste("trap", trapDat$trap, sep="_")
bci8$treeID <- paste("tree", bci8$treeID, sep="_")

# subset to one year and one species
bci8 <- subset(bci8, status=="A" & sp=="cordbi" & dbh>R50)
trapDat <- subset(trapDat, year=="2015" & SP6=="cordbi")

```
## Calculate euclidean distances
We will create a matrix with the rdist package to calculate the distance between each tree/trap combo. Will need to make this iterable when we move on to multiple years & sp.

```{r}
# Create a pairwise distance matrix, element [i (row), j(col)] is distance from tree i to trap j. 
# i.e. For each trap (col) you want to keep the tree (row) where the distance is within a range
dists <- cdist(bci8[,c("gx", "gy")], trapDat[,c("X", "Y")], metric="euclidean")

# Align back to bci8 to retain info
dists.df = as.data.frame(dists)
rownames(dists.df) <- unlist(bci8$treeID)
colnames(dists.df) <- unlist(trapDat$trap)
dists.df[1:5,1:5] #take a look

# merge back to bci8
bci8.dists = cbind(bci8, dists.df)

```
## Conspecific neighbourhood fecundity index
Lets start by calculating NFI, as it's the simpler option. 
```{r}
# function to calculate NFI
calculate_NFI <- function (trap) {
	subset(bci8.dists, bci8.dists[,trap]<=100) %>% # subset to only trees within a 100m radius of the trap
	group_by(treeID, year) %>% 
	mutate(a = dbh / eval(parse(text=trap))) %>% # divide dbh by dist
	ungroup() %>%
	dplyr::select(trap,"a") %>%
	summarise(trap = paste(trap), NFI = sum(a)) # sum over all trees to = 1 value per trap
}

# create list of trap IDs to pass through the function
traplist <- unique(trapDat[["trap"]])

# apply function to each trap
NFIdat <- lapply(traplist, calculate_NFI)

# bind output into one dataframe
NFIdat.b <- do.call(rbind, NFIdat)

# merge with trap data
trapConnect <- left_join(trapDat, NFIdat.b, by = "trap")
head(NFIdat.b) #take a look
```
## Hanski's Connectivity index
Now on to CI using the same format.

```{r}
# function to calculate HCI
calculate_CI <- function (trap) {
	subset(bci8.dists, bci8.dists[,trap]<=100) %>%
	group_by(treeID, year) %>% 
	mutate(a = dbh * exp( (-1/100) * eval(parse(text=trap)) ) ) %>% 
	ungroup() %>%
	dplyr::select(trap,"a") %>%
	summarise(trap = paste(trap), CI = sum(a))
}

CIdat <- lapply(list, calculate_CI)
CIdat.b <- do.call(rbind, CIdat)
trapConnect <- left_join(trapConnect, CIdat.b, by = "trap")
head(CIdat.b) #take a look
```

## Build and run some models

```{r}

m1 <- glm(cbind(abscised_seeds, viable_seeds) ~ 
	NFI, family = binomial(logit), 
	data = trapConnect)

m2 <- glm(cbind(abscised_seeds, viable_seeds) ~ 
	CI, family = binomial(logit), 
	data = trapConnect)

# compare
models <- list(m1, m2)
purrr::map_dfr(models, broom::tidy, 
							conf.int = TRUE, .id = "vars")

```
The model outputs look pretty similar, which is reassuring! I think Hanski's is probably the one to go for. Next steps are to include something to correct for edge effects and extend to multiple species and years.