---
title: "Connectivity"
author: "Eleanor Jackson"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: lumen
    highlight: pygments
---

I'm going to try two different measures of connectivity, __Hanski’s Connectivity Index (CI)__ [(Hanski 1994)](https://www.jstor.org/stable/5591) and the __Conspecific neighbourhood fecundity index (NFI)__ [(Jones & Comita 2008)](https://royalsocietypublishing.org/doi/10.1098/rspb.2008.0894#d3e405). 

They differ in that Hanski's CI includes an exponential function. CI also needs the average dispersal distance to scale your distances, which I'm just going to take as 50m for now.

I had to alter these equations a little. NFI uses the yearly mean total fruit set per tree to weight individuals - we don't have this, so I just used DBH, which assumes that larger individuals produce more fruit [(see here)](https://www.jstor.org/stable/2389015).

__Hanski's Connectivity Index__\

$CI = \sum exp(-\, \alpha \, dist_{ji}) A_j$ \

__Conspecific neighbourhood fecundity index__\

$NFI = \sum \frac {A_j}{dist_{ij}}$ \

_j_ = conspecific tree within a 100m radius of the trap\
_dist_ = distance between trap i and tree j\
$\alpha$ = 1/ avg. dispersal ability\
_A_ = DBH of tree j\

```{r, warning=FALSE, message=FALSE}
rm(list = ls())
library("tidyverse"); theme_set(theme_bw(base_size=10))
library("rdist")
library("lme4")
library("ggeffects")
library("knitr")
opts_chunk$set(tidy.opts=list(width.cutoff=50), tidy=FALSE, cache.lazy = FALSE,cache = TRUE)
```
## Read in and compile the data


```{r, warning=FALSE, results="hide"}
######## TREE DATA ########
# find and load all bci files in one go

# on home computer
file_names <- as.list(dir(
	path="/Users/eleanorjackson/OneDrive - University of Reading/SpatialPatterns/data/bci.tree", 
	pattern="bci.tree*"))
file_paths <- paste("../data/bci.tree/", file_names, sep="")

# on RACC
file_names <- as.list(dir(
	path="/home/users/ft840275/SpatialPatterns/data/bci.tree", 
	pattern="bci.tree*"))
file_paths <- paste("/home/users/ft840275/SpatialPatterns/data/bci.tree/", file_names, sep="")

lapply(file_paths, load, environment())

# expand the bci survey data so we can match it to nearest years of the trap data
# basically we are repeating each row of the bci datasets 5 times and assigning 2 years either side
e.bci.tree3 <- bci.tree3[rep(seq_len(nrow(bci.tree3)), each = 5), ]
e.bci.tree3$year <- rep(c("1988","1989","1990","1991","1992"))

e.bci.tree4 <- bci.tree4[rep(seq_len(nrow(bci.tree4)), each = 5), ]
e.bci.tree4$year <- rep(c("1993","1994","1995","1996","1997"))

e.bci.tree5 <- bci.tree5[rep(seq_len(nrow(bci.tree5)), each = 5), ]
e.bci.tree5$year <- rep(c("2002","2001","2000","1999","1998"))

e.bci.tree6 <- bci.tree6[rep(seq_len(nrow(bci.tree6)), each = 5), ]
e.bci.tree6$year <- rep(c("2003","2004","2005","2006","2007"))

e.bci.tree7 <- bci.tree7[rep(seq_len(nrow(bci.tree7)), each = 5), ]
e.bci.tree7$year <- rep(c("2008","2009","2010","2011","2012"))

e.bci.tree8 <- bci.tree8[rep(seq_len(nrow(bci.tree8)), each = 5), ]
e.bci.tree8$year <- rep(c("2013","2014","2015","2016","2017"))

# bind all the bci datasets together and add column for survey year
bind_rows(mget(ls(pattern="e.bci.tree*")), .id='df') %>%
    mutate(survey_year = case_when(df == "e.bci.tree8" ~ "2015",
                            df == "e.bci.tree7" ~ "2010",
                            df == "e.bci.tree6" ~ "2005",
                        	df == "e.bci.tree5" ~ "2000",
                        	df == "e.bci.tree4" ~ "1995",
                        	df == "e.bci.tree3" ~ "1990")) -> bci.bind

rm(list = ls(pattern='*bci.tree*')) # remove some of those big files

# read in other data
load("../data/bci.spptable.rdata") # species data
RST <- read.csv("../output/tables/R50.csv") # reproductive size threshold

# join bci with sp data and reproductive size threshold data
bci <- left_join(bci.bind, bci.spptable, by = "sp")
RST$sp <- tolower(RST$sp)
bci <- left_join(bci, RST, by = "sp")

# subset to only trees that are alive and are above reproductive size threshold
bci <- bci[!is.na(bci$R50), ]
bci <- dplyr::filter(bci, (bci[,"status"]=="A") & 
		(bci[,"dbh"]>bci[,"R50"]))

# add id tags to id values so they don't just look like numbers
bci$treeID <- paste("tree", bci$treeID, sep="_")

# will need year to be a character when we merge later
bci$year <- as.character(bci$year)

######## TRAP DATA ########
trapDat <- read.csv("../output/tables/proportionAbscisedPerTrap.csv") 
trapDat$trap <- paste("trap", trapDat$trap, sep="_")
trapDat$year <- as.character(trapDat$year)
# Subset to traps where there is more than 1 part per year per sp
trapDat <- subset(trapDat, sum_parts>5)

# save these so we don't have to do it every time
save(trapDat, bci, file = "cleanTrapandTree.RData")
load("cleanTrapandTree.RData")
```
## Calculate euclidean distances
We will create a pairwise matrix with the rdist package to calculate the distance between each tree/trap combo. 

```{r, warning=FALSE}

calculate_dist <- function (species, yr) {
	# subset tree data
	bd <- dplyr::filter(bci, (bci[,"status"]=="A") & (bci[,"sp"]==species) & 
		(bci[,"dbh"]>bci[,"R50"]) & (bci[,"year"]==yr))

	# subset trap data
	td <- dplyr::filter(trapDat, (trapDat[,"year"]==yr) & 
		(trapDat[,"SP6"]==species))

	# create distance matrix using x and y co-ordinates
	dists <- cdist(bd[,c("gx", "gy")], td[,c("X", "Y")], metric="euclidean")
	dists.df <- as.data.frame(dists)

	# label col and row names with trap and tree IDs
	rownames(dists.df) <- unlist(bd$treeID)
	colnames(dists.df) <- unlist(td$trap)

	# bind with the bci data
	cbind(bd, dists.df)
}

# create lists of sp and year to pass to function
#sp.list <- unique(bci[["sp"]])
sp.list <- c("cordbi", "mourmy")
year.list <- unique(bci[["year"]]) 

# apply function to all pairwise combinations of year and species
all.dists <- outer(sp.list, year.list, FUN = Vectorize(calculate_dist))
bci.dists <- bind_rows(all.dists)

#or just apply function to one year and one species
#bci.dists <- calculate_dist("cordbi", "2015")

```
## Conspecific neighbourhood fecundity index
Lets start by calculating NFI.
```{r, warning=FALSE}
# function to calculate NFI
calculate_NFI <- function (trap) {
	dplyr::filter(bci.dists, bci.dists[,trap] <= 100) %>% # subset to only trees within a 100m radius of the trap
	group_by(treeID, year, sp) %>% 
	mutate(a = dbh / eval(parse(text=trap))) %>% # divide dbh by dist
	ungroup() %>%
	dplyr::select(year, sp, trap, a) %>%
	group_by(year, sp) %>%
	summarise(trap = paste(trap), NFI = sum(a), .groups = "drop") # sum over all trees to = 1 value per trap
}

# create list of trap IDs to pass through the function
trap.list <- colnames(select(bci.dists, matches("trap_")))

# apply function to each trap and parallelize
numCores <- detectCores()
NFIdat <- mclapply(trap.list, calculate_NFI, mc.cores = numCores)

# bind output into one dataframe
NFIdat.b <- bind_rows(NFIdat)

# merge with trap data
trapConnect <- left_join(NFIdat.b, trapDat,  
	by = c("trap", "year", "sp" = "SP6"))

head(NFIdat.b) #take a look at NFI values
```
## Hanski's Connectivity index
Now on to CI using the same format.

```{r, warning=FALSE}
# function to calculate CI
calculate_CI <- function (trap) {
	dplyr::filter(bci.dists, bci.dists[,trap] <= 100) %>% 
	group_by(treeID, year, sp) %>% 
	mutate(a = dbh * exp( (-1/50) * eval(parse(text=trap)) ) ) %>% 
	ungroup() %>%
	dplyr::select(year, sp, trap, a) %>%
	group_by(year, sp) %>%
	summarise(trap = paste(trap), CI = sum(a), .groups = "drop")
}

CIdat <- mclapply(trap.list, calculate_CI, mc.cores = numCores)
CIdat.b <- bind_rows(CIdat)
trapConnect <- left_join(trapConnect, CIdat.b, by = c("trap", "year", "sp"))

head(CIdat.b) #take a look at CI values
```

## Lets see if NFI and CI correlate
```{r, warning=FALSE}
ggplot(trapConnect) +
	geom_point(aes(x= CI, y= NFI))

# save these so we don't have to run this every time
save(trapConnect, file = "trapConnect.RData")
```